\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-08-07 analysis/paper/paper.Rmd
fix figure 7

```{r baselines, echo=FALSE, eval=eval_cs3}
cox_ci <- ggplot(baseCoxComparisonData, aes(Time)) +
  geom_step(aes(y = APS, colour = "APS")) +
  geom_step(aes(y = Full, col = "Full")) +
  geom_step(aes(y = SPS, colour = "SPS")) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of Death", x = "Follow-up time (Years)", color = "Models") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme+ scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=8)
```

@@

abCB <- data.frame(Time = abCbFull[, 1],
                   absRisk = rowMeans(abCbFull[, -c(1)]))
abCB$Model <- "Penalized estimation"

@@

```{r cs3Final, echo=FALSE, eval=eval_cs3}
reg_ci <- ggplot(ab, aes(Time)) +
  geom_step(data = baseCoxComparisonData, mapping = aes(x = Time,y = Full, colour = "Cox")) +
  geom_line(aes(y = absRisk, colour = Model)) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of death", x = "Follow-up time (years)", color = "Models") +
  # ggtitle("Absolute risk curves") +
  scale_color_manual(values = q4, breaks = c("K-M","Cox","Penalized Cox","Penalized CB")) + 
  paper_gg_theme+ scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=9)
```

@@

  # if (missing(cause)) stop("Argument cause should be the event type for which we predict the absolute risk.")
  # the output of absoluteRisk is an array with dimension dependening on the length of the requested times:
  # case 1: the number of time points is 1
  #         dim(array) =  (length(time), NROW(newdata), number of causes in the data)
  if (length(times) == 1) {
    a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
    p <- matrix(a, ncol = 1)
  } else {
    # case 2 a) zero is included in the number of time points
    if (0 %in% times) {
      # dim(array) =  (length(time)+1, NROW(newdata)+1, number of causes in the data)
      a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
      p <- t(a)
    } else {
      # case 2 b) zero is not included in the number of time points (but the absoluteRisk function adds it)
      a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
      ### we need to invert the plot because, by default, we get cumulative incidence
      #a[, -c(1)] <- 1 - a[, -c(1)]
      ### we remove time 0 for everyone, and remove the time column
      a <- a[-c(1), -c(1)] ### a[-c(1), ] to keep times column, but remove time 0 probabilities
      # now we transpose the matrix because in riskRegression we work with number of
      # observations in rows and time points in columns
      p <- t(a)
    }
  }
  if (NROW(p) != NROW(newdata) || NCOL(p) != length(times)) {
    stop(paste("\nPrediction matrix has wrong dimensions:\nRequested newdata x times: ", 
               NROW(newdata), " x ", length(times), "\nProvided prediction matrix: ", 
               NROW(p), " x ", NCOL(p), "\n\n", sep = ""))
  }
  p
}
```

@@

data_brier <- bind_rows(
  brierFinalResults$Brier$score,
  filter(brierFull$Brier$score, model == "PredictionModel")
  ) %>% 
  mutate(model = as.character(model), # First return to string, so types match
         model = case_when(
    model == "Null model" ~ "K-M",
    model == "PredictionModel" ~ "Cox",
    TRUE ~ model
  ),
  model = factor(model, levels = c("Cox", "Penalized Cox", "Penalized Logistic", "K-M"),
                 labels = c("Cox", "Pen. Cox", "Pen. Logistic", "K-M")))

reg_brier <- ggplot(data = data_brier, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Follow-up time (years)") +
  ylab("Brier score") +
  # ggtitle("Brier score over time") +
  labs(color = "Models") +
  paper_gg_theme +  
  scale_color_manual(values = q4) + scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=8)

>>>>

```{r baselines, echo=FALSE, eval=eval_cs3}
cox_ci <- ggplot(baseCoxComparisonData, aes(Time)) +
  geom_step(aes(y = APS, colour = "APS")) +
  geom_step(aes(y = Full, col = "Full")) +
  geom_step(aes(y = SPS, colour = "SPS")) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of Death", x = "Follow-up time (Years)", color = "Models") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme + 
  scale_x_continuous(labels = scaleFUNx, breaks = round(seq(0, 5, by = 1))) + 
  scale_y_continuous(labels = scaleFUNy, n.breaks = 8)
```

@@

abCB <- data.frame(Time = abCbFull[, 1],
                   absRisk = rowMeans(abCbFull[, -c(1)]))
abCB$Model <- "Penalized CB"

@@

```{r cs3Final, echo=FALSE, eval=eval_cs3}
reg_ci <- ggplot(ab, aes(Time)) +
  geom_step(data = baseCoxComparisonData, mapping = aes(x = Time,y = Full, colour = "Cox")) +
  geom_line(aes(y = absRisk, colour = Model)) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of death", x = "Follow-up time (years)", color = "Models") +
  # ggtitle("Absolute risk curves") +
  scale_color_manual(values = q4, breaks = c("K-M", "Cox", "Penalized Cox", "Penalized CB")) + 
  paper_gg_theme + 
  scale_x_continuous(labels = scaleFUNx, breaks = round(seq(0, 5, by = 1))) + 
  scale_y_continuous(labels = scaleFUNy, n.breaks = 9)
```

@@

  # if (missing(cause)) stop("Argument cause should be the event type for which we predict the absolute risk.")
  # the output of absoluteRisk is an array with dimension depending on the length of the requested times:
  # case 1: the number of time points is 1
  #         dim(array) =  (length(time), NROW(newdata), number of causes in the data)
  if (length(times) == 1) {
    a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
    p <- matrix(a, ncol = 1)
  } else {
    # case 2 a) zero is included in the number of time points
    if (0 %in% times) {
      # dim(array) =  (length(time)+1, NROW(newdata)+1, number of causes in the data)
      a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
      p <- t(a)
    } else {
      # case 2 b) zero is not included in the number of time points (but the absoluteRisk function adds it)
      a <- casebase::absoluteRisk(object, newdata = newdata, time = times)
      ### we need to invert the plot because, by default, we get cumulative incidence
      #a[, -c(1)] <- 1 - a[, -c(1)]
      ### we remove time 0 for everyone, and remove the time column
      a <- a[-c(1), -c(1)] ### a[-c(1), ] to keep times column, but remove time 0 probabilities
      # now we transpose the matrix because in riskRegression we work with number of
      # observations in rows and time points in columns
      p <- t(a)
    }
  }
  if (NROW(p) != NROW(newdata) || NCOL(p) != length(times)) {
    stop(paste("\nPrediction matrix has wrong dimensions:\nRequested newdata x times: ", 
               NROW(newdata), " x ", length(times), "\nProvided prediction matrix: ", 
               NROW(p), " x ", NCOL(p), "\n\n", sep = ""))
  }
  p
}
```

@@

data_brier <- bind_rows(
  brierFinalResults$Brier$score,
  filter(brierFull$Brier$score, model == "PredictionModel")
  ) %>% 
  mutate(model = as.character(model), # First return to string, so types match
         model = case_when(
    model == "Null model" ~ "K-M",
    model == "PredictionModel" ~ "Cox",
    TRUE ~ model
  ),
  model = factor(model, levels = c("Cox", "Penalized Cox", "Penalized CB", "K-M"),
                 labels = c("Cox", "Pen. Cox", "Pen. CB", "K-M")))

reg_brier <- ggplot(data = data_brier, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Follow-up time (years)") +
  ylab("Brier score") +
  # ggtitle("Brier score over time") +
  labs(color = "Models") +
  paper_gg_theme +  
  scale_color_manual(values = q4) + 
  scale_x_continuous(labels = scaleFUNx, 
                     breaks = round(seq(0, 5, by = 1))) + 
  scale_y_continuous(labels = scaleFUNy, n.breaks = 8)

### 2020-08-07 analysis/paper/paper.tex
fix figure 7

This report was generated on 2020-08-06 17:31:23 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 4.0.2 (2020-06-22)
#> Platform: x86_64-apple-darwin17.0 (64-bit)
#> Running under: macOS Catalina 10.15.6
#> 
#> Matrix products: default
#> BLAS:   /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRblas.dylib
#> LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib
#> 
#> locale:
#> [1] en_CA.UTF-8/en_CA.UTF-8/en_CA.UTF-8/C/en_CA.UTF-8/en_CA.UTF-8
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.6               dotwhisker_0.5.0         
#>  [3] timereg_1.9.4             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.0              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.0.2               ggplot2_3.3.2            
#> [13] tidyverse_1.3.0           lubridate_1.7.8          
#> [15] visreg_2.6.1              tibble_3.0.3             
#> [17] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.1-12          
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_3.0-2             
#> [25] Matrix_1.2-18             data.table_1.12.8        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0.9000      
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-2          ellipsis_0.3.1     
#>  [4] ggstance_0.3.4      htmlTable_1.13.3    base64enc_0.1-3    
#>  [7] fs_1.4.1            rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.1-0      
#> [13] xml2_1.3.2          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.6.1      cluster_2.1.0      
#> [19] dbplyr_1.4.3        png_0.1-7           compiler_4.0.2     
#> [22] httr_1.4.1          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.55       tools_4.0.2         gtable_0.3.0       
#> [31] Rcpp_1.0.5          cellranger_1.1.0    vctrs_0.3.2        
#> [34] nlme_3.1-148        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.5         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.6       zoo_1.8-8          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.14        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.7          shape_1.4.4        
#> [58] rlang_0.4.7         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-41     labeling_0.3        htmlwidgets_1.5.1  
#> [64] cmprsk_2.2-9        tidyselect_1.1.0    deSolve_1.28       
#> [67] R6_2.4.1            generics_0.0.2      Hmisc_4.4-0        
#> [70] multcomp_1.4-13     DBI_1.1.0           withr_2.2.0        
#> [73] pillar_1.4.6        haven_2.2.0         foreign_0.8-80     
#> [76] mgcv_1.8-31         nnet_7.3-14         mstate_0.2.12      
#> [79] modelr_0.1.6        crayon_1.3.4        rmarkdown_2.1      
#> [82] jpeg_0.1-8.1        grid_4.0.2          readxl_1.3.1       
#> [85] reprex_0.3.0        digest_0.6.25       webshot_0.5.2      
#> [88] numDeriv_2016.8-1.1 stats4_4.0.2        munsell_0.5.0      
#> [91] viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    Jesse-Islam-master /Users/maxturgeon/Documents/git_repos/cbpaper
#> Head:     [32e6bf4] 2020-08-05: Merge pull request #25 from sahirbhatnagar/editMax
\end{CodeOutput}
\end{CodeChunk}

>>>>

This report was generated on 2020-08-07 09:30:39 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 4.0.2 (2020-06-22)
#> Platform: x86_64-apple-darwin17.0 (64-bit)
#> Running under: macOS Catalina 10.15.6
#> 
#> Matrix products: default
#> BLAS:   /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRblas.dylib
#> LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib
#> 
#> locale:
#> [1] en_CA.UTF-8/en_CA.UTF-8/en_CA.UTF-8/C/en_CA.UTF-8/en_CA.UTF-8
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.7.0               dotwhisker_0.5.0         
#>  [3] timereg_1.9.4             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.1              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.1.1               ggplot2_3.3.2            
#> [13] tidyverse_1.3.0           lubridate_1.7.9          
#> [15] visreg_2.7.0              tibble_3.0.3             
#> [17] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.2-3           
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_4.0-2             
#> [25] Matrix_1.2-18             data.table_1.13.0        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0.9000      
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] ggstance_0.3.4      htmlTable_1.13.3    base64enc_0.1-3    
#>  [7] fs_1.4.1            rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.1-1      
#> [13] xml2_1.3.2          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.7.0      cluster_2.1.0      
#> [19] dbplyr_1.4.4        png_0.1-7           compiler_4.0.2     
#> [22] httr_1.4.2          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.55       tools_4.0.2         gtable_0.3.0       
#> [31] Rcpp_1.0.5          cellranger_1.1.0    vctrs_0.3.2        
#> [34] nlme_3.1-148        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.6         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.6       zoo_1.8-8          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.14        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.7          shape_1.4.4        
#> [58] rlang_0.4.7         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-41     labeling_0.3        htmlwidgets_1.5.1  
#> [64] cmprsk_2.2-10       tidyselect_1.1.0    deSolve_1.28       
#> [67] plyr_1.8.6          R6_2.4.1            generics_0.0.2     
#> [70] Hmisc_4.4-0         multcomp_1.4-13     DBI_1.1.0          
#> [73] withr_2.2.0         pillar_1.4.6        haven_2.3.1        
#> [76] foreign_0.8-80      mgcv_1.8-31         nnet_7.3-14        
#> [79] mstate_0.2.12       modelr_0.1.8        crayon_1.3.4       
#> [82] rmarkdown_2.3       jpeg_0.1-8.1        grid_4.0.2         
#> [85] readxl_1.3.1        blob_1.2.1          reprex_0.3.0       
#> [88] digest_0.6.25       webshot_0.5.2       numDeriv_2016.8-1.1
#> [91] stats4_4.0.2        munsell_0.5.0       viridisLite_0.3.0  
#> [94] quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master /Users/maxturgeon/Documents/git_repos/cbpaper
#> Remote:   master @ origin (git@github.com:sahirbhatnagar/cbpaper.git)
#> Head:     [06b07b3] 2020-08-06: incorporate Ollis comments on CS 3
\end{CodeOutput}
\end{CodeChunk}
